---
title: "SML_Final_project"
author: "Soo"
date: "6/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

First include the relevant libraries.

```{r,warning=FALSE}
library(pixmap)
```

# Picture Extraction

Use basic regular expressions to extract only the pictures.

```{r,warning=FALSE}

setwd("/Users/LenaShin/Desktop/")

# Get directory structure
dir_list = dir(path="CroppedYale/",all.files=FALSE,recursive=TRUE)

# Update directory char vector by removing cases
dir_list_new <- dir_list[-grep("WS_FTP|Ambient|info|DEADJOE|bad",x=dir_list)]
```

Using character string **dir_list_new**, create a function that extracts all the face pictures of a person, vectorizes it, and outputs a matix with the features in the columns and samples in the rows.

```{r,warning=FALSE}

pictures <- function(folder_name,string,dir_list,n.col=1){

  index <- grep(string,x=dir_list)
  face_matrix <- matrix(nrow=length(index),ncol=n.col)
  for(i in 1:length(dir_list[index])){

    face <- read.pnm(file = paste(folder_name,x=dir_list[index][i],sep=""))
    face_matrix[i,] <- c(getChannels(face))
    }
  
  return(face_matrix)
}

```

Likewise, create a function that extracts n.pic number of sample pictures of a person and outputs a matrix of the pictures.

```{r,warning=FALSE}

sample_pictures <- function(folder_name,string,dir_list,n.col=1,n.pic=1){

  index <- grep(string,x=dir_list)
  face_matrix <- matrix(nrow=n.pic,ncol=n.col)
  random_index <- sample(index,size=n.pic,replace=FALSE)
  for(i in 1:n.pic){
    face <- read.pnm(file = paste(folder_name,x=dir_list[random_index][i],sep=""))
    face_matrix[i,] <- c(getChannels(face))
    }
  return(face_matrix)
}

```

# Exploratory Data Analysis

Store all the pictures of the first person.

```{r,warning=FALSE}

setwd("/Users/LenaShin/Desktop/Summer2019/SML")
person_1 <- pictures("CroppedYale/","yaleB01",dir_list_new,n.col=32256)

# Print the first 3 pictures of the person
par(mfrow=c(1,1), mai=c(0.1,0.1,0.1,0.1))
plot(pixmapGrey(data=c(person_1[1,],person_1[2,],person_1[3,]),nrow=192,ncol=168*3,cellres=1))

```

Get the mean face of the first person and plot it.

```{r}
par(mfrow=c(1,1),mai=c(1,1,1,1))
mean_face_1<- apply(person_1, 2, mean)
plot(pixmapGrey(data=mean_face_1,nrow=192,ncol=168,cellres=1))
```

Using the function, extract all the pictures of all the people. (It takes much time to contain all the pictures into a matrix. Might not want to do it..)

```{r,warning=FALSE}

# Get the list of the folder names of each peoson
people_list <- vector(mode="character",length=39)
for(i in 1:39){
  ifelse(i<10, people_list[i] <- paste("yaleB0",i,sep=""), people_list[i] <- paste("yaleB",i,sep=""))
}

# Note that there is no folder named "yaleB14". So, remove it. Then, there are 38 people in total.
people_list <- people_list[-14]


# To create a matrix to contain all the pictures of all the people, run the code below

#setwd("/Users/LenaShin/Desktop/")
#people <- person1
#for(i in 2:38){
#  people <- rbind(people,pictures("CroppedYale/",people_list[i],dir_list_new,n.col=32256))
#}
```

Get the mean face of all the people and plot it.

```{r}
#par(mfrow=c(1,1),mai=c(1,1,1,1))
#mean_face<- apply(people, 2, mean)
#plot(pixmapGrey(data=mean_face,nrow=192,ncol=168,cellres=1))
```

Since there are too many pictures of the people, let's select 3 (or more as you want) sample pictures for each person.

```{r,warning=FALSE}

setwd("/Users/LenaShin/Desktop/Summer2019/SML")

# Create a matrix to contain the sample pictures of the 38 people

people_sample <- sample_pictures("CroppedYale/",people_list[1],dir_list_new,n.col=32256,n.pic=10)

for(i in 2:38){
  people_sample <- rbind(people_sample,sample_pictures("CroppedYale/",people_list[i],dir_list_new,n.col=32256,n.pic=3))
}

```

Get the mean face from the random pictures of the 38 people and plot it.

```{r,warning=FALSE}
par(mfrow=c(1,1),mai=c(1,1,1,1))
# We need to remove the last column which is the sample index
mean_face_sample <- apply(people_sample, 2, mean)
plot(pixmapGrey(data=mean_face_sample,nrow=192,ncol=168,cellres=1))
```

Now, run PCA on all the pictures of the first person without scaling, and see how many principal components are important.

```{r,warning=FALSE}

person_1_pca <- prcomp(person_1)
plot(seq(from=1,to=64,by=1),(person_1_pca$sdev)^2/sum((person_1_pca$sdev)^2),type="l",xlab="number of PCs",ylab="ratio of the variance")
points(seq(from=1,to=64,by=1),(person_1_pca$sdev)^2/sum((person_1_pca$sdev)^2))

```

The first three principal components explain about 90% of the variance in the data set. Now, let's visualize the first five components for comparison.

```{r,warning=FALSE}

par(mfrow=c(1,1), mai=c(0.1,0.1,0.1,0.1))
plot(pixmapGrey(data=c(person_1_pca$rotation[,1],person_1_pca$rotation[,2],person_1_pca$rotation[,3],person_1_pca$rotation[,4],person_1_pca$rotation[,5]),nrow=192,ncol=168*5,cellres=1))

```

Let's scale it and do the same thing.

```{r,warning=FALSE}

person_1_scaled <- scale(person_1,center=TRUE, scale=FALSE)
person_1_scaled.pca <- prcomp(person_1_scaled)
plot(seq(from=1,to=64,by=1),(person_1_scaled.pca$sdev)^2/sum((person_1_scaled.pca$sdev)^2),type="l",xlab="number of PCs",ylab="ratio of the variance")
points(seq(from=1,to=64,by=1),(person_1_scaled.pca$sdev)^2/sum((person_1_scaled.pca$sdev)^2))
par(mfrow=c(1,1), mai=c(0.1,0.1,0.1,0.1))
plot(pixmapGrey(data=c(person_1_scaled.pca$rotation[,1],person_1_scaled.pca$rotation[,2],person_1_scaled.pca$rotation[,3],person_1_scaled.pca$rotation[,4],person_1_scaled.pca$rotation[,5]),nrow=192,ncol=168*5,cellres=1))

```

It seems like scaling in this case didn't make any difference in terms of the variance of each PCs.

Now, let's extract the pictures of three different people and compare the projection on the first two PCs. 

```{r,warning=FALSE}
setwd("/Users/LenaShin/Desktop/Summer2019/SML")
three_people <- person_1

n<-vector(mode="numeric",length=3)
for(i in 2:3){
  three_people <- rbind(three_people,pictures("CroppedYale/",people_list[i],dir_list_new,n.col=32256))
  n[i] <- dim(three_people)[1]-sum(n)
}

three_people_pca <- prcomp(three_people)

plot(seq(from=1,to=64*3,by=1),(three_people_pca$sdev)^2/sum((three_people_pca$sdev)^2),type="l", xlab="number of PCs", ylab="ratio of the variance")
points(seq(from=1,to=64*3,by=1),(three_people_pca$sdev)^2/sum((three_people_pca$sdev)^2))

```

Let's organize the data by adding a column that indicates which person the row refers to. I'll keep the index of 10 pictures of each person as a test set in case.

```{r,warning=FALSE}

type <- as.matrix(c(rep("1",times=n[1]),rep("2",times=n[2]),rep("3",times=n[3])),nrow=sum(n),ncol=1)

three_people_pca$x <- cbind(three_people_pca$x,type)
dim(three_people_pca$x)
colnames(three_people_pca$x) <- c(colnames(three_people_pca$x[,-193]),"type")
colnames(three_people_pca$x)

```


```{r,warning=FALSE}
plot(pixmapGrey(data=c(three_people_pca$rotation[,1],three_people_pca$rotation[,2],three_people_pca$rotation[,3],three_people_pca$rotation[,4],three_people_pca$rotation[,5]),nrow=192,ncol=168*5,cellres=1))
```

```{r,warning=FALSE}
par(mfrow=c(1,1), mai=c(1, 1, 1, 1))
plot(three_people_pca$x[,1], three_people_pca$x[,2],pch=16, xlab="First Principle Component", ylab="Second Principle Component" )
```
